---
title: "TafelTrainer Data Analysis Across Levels"
---

```{r setup, include=FALSE}
# Load required libraries
library(skimr)
library(tidyverse)
```

### Description
This analysis prepares data from Levels 1, 2, and 3 of the TafelTrainer dataset. The goal is to create datasets for GraafTel input (first, middle, and last encounters) and conduct descriptive analyses.

### Data Preparation
```{r}
# Load dataset
dat <- readRDS('./TafelTrainer pilot data/tt_responses_all_clean.rds')

# Initial filtering: Ensure 'correct' is numeric and clean `cue_text`
dat <- dat %>%
  mutate(
    correct = as.numeric(correct),
    cue_text = str_replace_all(cue_text, "\\+", " ") # Replace "+" with a space
  )

# Define valid responses (numeric values ≤ 100)
valid_responses <- 0:100

# Filter for valid 'given_response'
dat <- dat %>%
  filter(as.numeric(given_response) %in% valid_responses)

# Add multiplier and multiplicand columns globally to the dataset
dat <- dat %>%
  mutate(
    multiplier = as.numeric(str_extract(cue_text, "^\\d+")), # Extract the first number in `cue_text`
    multiplicand = as.numeric(str_extract(cue_text, "\\d+$")) # Extract the second number in `cue_text`
  )

# Function to process each level and add encounter counter
prepare_level_data <- function(data, level_filter) {
  data %>%
    filter(level == level_filter) %>%
    arrange(user_id, cue_text, session_id) %>%
    group_by(user_id, cue_text) %>%
    mutate(encounter_num = row_number()) %>%
    ungroup()
}

# Prepare full datasets for Levels 1, 2, and 3 (including `encounter_num`)
dat_level1 <- prepare_level_data(dat, 1)
dat_level2 <- prepare_level_data(dat, 2)
dat_level3 <- prepare_level_data(dat, 3)

# Function to filter specific encounters dynamically
filter_encounter <- function(data, encounter_type) {
  data %>%
    group_by(user_id, cue_text) %>%
    filter(
      case_when(
        # First encounter
        encounter_type == "first" ~ encounter_num == 1,
        # Middle encounter: Use ceiling(n() / 2), but if it results in 1, use 2 if available
        encounter_type == "middle" ~ {
          middle_encounter <- ceiling(n() / 2)
          encounter_num == if_else(middle_encounter == 1 & n() > 1, 2L, middle_encounter)
        },
        # Last encounter
        encounter_type == "last" ~ encounter_num == n(),
        # Default: Exclude rows
        TRUE ~ FALSE
      )
    ) %>%
    ungroup()
}

# Generate encounter-specific datasets dynamically from the full datasets
dat_level1_first <- filter_encounter(dat_level1, "first")
dat_level1_middle <- filter_encounter(dat_level1, "middle")
dat_level1_last <- filter_encounter(dat_level1, "last")

dat_level2_first <- filter_encounter(dat_level2, "first")
dat_level2_middle <- filter_encounter(dat_level2, "middle")
dat_level2_last <- filter_encounter(dat_level2, "last")

dat_level3_first <- filter_encounter(dat_level3, "first")
dat_level3_middle <- filter_encounter(dat_level3, "middle")
dat_level3_last <- filter_encounter(dat_level3, "last")
```

### Export Data for GraafTel
```{r}
# Function to export datasets
export_graaftel_data <- function(data, level, encounter_type) {
  # Construct the filename
  file_name <- paste0("./graaftel_input_level", level, "_", encounter_type, ".csv")
  # Select the required columns
  data_to_export <- data %>%
    select(user_id, cue_text, correct)
  # Write data to CSV (without column headers)
  write.table(data_to_export, file_name, row.names = FALSE, col.names = FALSE, sep = ",")
}

# Export data for Level 1
export_graaftel_data(dat_level1_first, 1, "first")
export_graaftel_data(dat_level1_middle, 1, "middle")
export_graaftel_data(dat_level1_last, 1, "last")

# Export data for Level 2
export_graaftel_data(dat_level2_first, 2, "first")
export_graaftel_data(dat_level2_middle, 2, "middle")
export_graaftel_data(dat_level2_last, 2, "last")

# Export data for Level 3
export_graaftel_data(dat_level3_first, 3, "first")
export_graaftel_data(dat_level3_middle, 3, "middle")
export_graaftel_data(dat_level3_last, 3, "last")
```

### Verify Outputs
```{r}
# Prepare data for verification
# Combine all levels and encounter types to verify overall distribution
verify_data <- bind_rows(
  dat_level1 %>% mutate(level == 1),
  dat_level2 %>% mutate(level == 2),
  dat_level3 %>% mutate(level == 3)
)

# Plot histogram for the distribution of responses
# Ensures that the data used for GraafTel matches expected values
verify_data %>%
  select(given_response, cue_text, level) %>%
  filter(given_response %in% 0:100) %>%
  mutate(given_response = as.numeric(given_response)) %>%
  ggplot(aes(x = given_response, fill = factor(level))) +
  geom_histogram(binwidth = 1, position = "dodge") +
  theme_minimal() +
  labs(
    title = "Distribution of Responses Across Levels",
    x = "Given Response",
    y = "Count",
    fill = "Level"
  ) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_brewer(palette = "Set3")

# Calculate mean accuracy for each fact_id and level
mean_accuracy_data <- verify_data %>%
  group_by(level, cue_text) %>% # Group by level and fact_id
  summarize(mean_accuracy = mean(correct), .groups = "drop") # Calculate mean accuracy

# Create a box plot of mean accuracies
mean_accuracy_data %>%
  ggplot(aes(x = factor(level), y = mean_accuracy, fill = factor(level))) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Distribution of Mean Accuracies by Level",
    x = "Level",
    y = "Mean Accuracy",
    fill = "Level"
  ) +
  scale_fill_brewer(palette = "Set3") +
  theme(legend.position = "none")
```
### Descriptive stats
## Modified code with reusability
```{r}
generate_descriptive_data <- function(data) {
  # Group by cue_text and compute summary statistics
  data %>%
    select(given_response, cue_text, correct, level, fact_id) %>%
    mutate(cue_text=fct_reorder(cue_text,fact_id)) %>% 
    group_by(cue_text) %>%
    summarise(mean_acc = mean(correct), n=n(), se = sd(correct)/sqrt(n)
    ) %>%
    mutate(
      multiplier = as.numeric(str_extract(cue_text, "^\\d+")), # Extract the first number in `cue_text`
      multiplicand = as.numeric(str_extract(cue_text, "\\d+$")) # Extract the second number in `cue_text`
    ) %>%
    mutate(cue_text=fct_reorder(cue_text,multiplicand))
}


# Generate descriptive data
dat4plot_level1 <- generate_descriptive_data(dat_level1)
dat4plot_level1_first <- generate_descriptive_data(dat_level1_first)
dat4plot_level1_middle <- generate_descriptive_data(dat_level1_middle)
dat4plot_level1_last <- generate_descriptive_data(dat_level1_last)
dat4plot_level2 <- generate_descriptive_data(dat_level2)
dat4plot_level2_first <- generate_descriptive_data(dat_level2_first)
dat4plot_level2_middle <- generate_descriptive_data(dat_level2_middle)
dat4plot_level2_last <- generate_descriptive_data(dat_level2_last)
dat4plot_level3 <- generate_descriptive_data(dat_level3)
dat4plot_level3_first <- generate_descriptive_data(dat_level3_first)
dat4plot_level3_middle <- generate_descriptive_data(dat_level3_middle)
dat4plot_level3_last <- generate_descriptive_data(dat_level3_last)


# Function to plot heatmaps with properly ordered and labeled axes
plot_heatmap <- function(data, title) {
  # Update multiplier to numeric for proper ordering and labeling
  data <- data %>%
    mutate(multiplier = as.numeric(gsub("col", "", multiplier)))

  data %>%
    ggplot(aes(x = as.factor(multiplier), y = as.factor(multiplicand), fill = mean_acc, label = round(mean_acc, 2))) +
    geom_tile() +
    geom_text(size = 3) + # Adjust text size for readability
    scale_fill_gradient2(
      limits = c(0.7, 1),
      midpoint = 0.85,
      low = 'blue',
      mid = "red",
      high = 'yellow',
      guide = 'colorbar',
      aesthetics = 'fill',
      breaks = seq(0.7, 1, by = 0.05) # Reduce the number of breaks for readability
    ) +
    theme_minimal() +
    labs(
      title = title,
      x = "Multiplier",
      y = "Multiplicand",
      fill = "Mean Accuracy"
    ) +
    theme(
      axis.text.x = element_text(angle = 0), # Ensure x-axis labels are horizontal
      axis.text.y = element_text(angle = 0), # Ensure y-axis labels are clear
      legend.key.height = unit(0.5, "cm") # Adjust legend key size for readability
    )
}

# Function to plot bar plots with error bars
plot_barplot <- function(data, title) {
  data %>%
    ggplot(aes(
      x = cue_text, y = mean_acc, ymax = mean_acc + se, ymin = mean_acc - se, fill = as.factor(multiplicand)
    )) +
    geom_bar(stat = 'identity') +
    geom_errorbar() +
    scale_x_discrete(
      breaks = levels(data$cue_text)[seq(1, length(levels(data$cue_text)), by = 10)] # Show every 10th label correctly
    ) +
    theme_minimal() +
    labs(
      title = title,
      x = "Cue Text (Multiplication Fact)",
      y = "Mean Accuracy",
      fill = "Multiplicand"
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = 'top')
}

# Generate plots for each level and overall
# Heatmaps
heatmap_level1 <- plot_heatmap(dat4plot_level1, "Heatmap of Mean Accuracy (Level 1: Overall)")
heatmap_level1_first <- plot_heatmap(dat4plot_level1_first, "Heatmap of Mean Accuracy (Level 1: First)")
heatmap_level1_middle <- plot_heatmap(dat4plot_level1_middle, "Heatmap of Mean Accuracy (Level 1: Middle)")
heatmap_level1_last <- plot_heatmap(dat4plot_level1_last, "Heatmap of Mean Accuracy (Level 1: Last)")
heatmap_level2 <- plot_heatmap(dat4plot_level2, "Heatmap of Mean Accuracy (Level 2: Overall)")
heatmap_level2_first <- plot_heatmap(dat4plot_level2_first, "Heatmap of Mean Accuracy (Level 2: First)")
heatmap_level2_middle <- plot_heatmap(dat4plot_level2_middle, "Heatmap of Mean Accuracy (Level 2: Middle)")
heatmap_level2_last <- plot_heatmap(dat4plot_level2_last, "Heatmap of Mean Accuracy (Level 2: Last)")
heatmap_level3 <- plot_heatmap(dat4plot_level3, "Heatmap of Mean Accuracy (Level 3: Overall)")
heatmap_level3_first <- plot_heatmap(dat4plot_level3_first, "Heatmap of Mean Accuracy (Level 3: First)")
heatmap_level3_middle <- plot_heatmap(dat4plot_level3_middle, "Heatmap of Mean Accuracy (Level 3: Middle)")
heatmap_level3_last <- plot_heatmap(dat4plot_level3_last, "Heatmap of Mean Accuracy (Level 3: Last)")

# Bar Plots
barplot_level1 <- plot_barplot(dat4plot_level1, "Bar Plot of Mean Accuracy (Level 1: Overall)")
barplot_level1_first <- plot_barplot(dat4plot_level1_first, "Bar Plot of Mean Accuracy (Level 1: First)")
barplot_level1_middle <- plot_barplot(dat4plot_level1_middle, "Bar Plot of Mean Accuracy (Level 1: Middle)")
barplot_level1_last <- plot_barplot(dat4plot_level1_last, "Bar Plot of Mean Accuracy (Level 1: Last)")
barplot_level2 <- plot_barplot(dat4plot_level2, "Bar Plot of Mean Accuracy (Level 2: Overall)")
barplot_level2_first <- plot_barplot(dat4plot_level2_first, "Bar Plot of Mean Accuracy (Level 2: First)")
barplot_level2_middle <- plot_barplot(dat4plot_level2_middle, "Bar Plot of Mean Accuracy (Level 2: Middle)")
barplot_level2_last <- plot_barplot(dat4plot_level2_last, "Bar Plot of Mean Accuracy (Level 2: Last)")
barplot_level3 <- plot_barplot(dat4plot_level3, "Bar Plot of Mean Accuracy (Level 3: Overall)")
barplot_level3_first <- plot_barplot(dat4plot_level3_first, "Bar Plot of Mean Accuracy (Level 3: First)")
barplot_level3_middle <- plot_barplot(dat4plot_level3_middle, "Bar Plot of Mean Accuracy (Level 3: Middle)")
barplot_level3_last <- plot_barplot(dat4plot_level3_last, "Bar Plot of Mean Accuracy (Level 3: Last)")

# Display plots
heatmap_level1
heatmap_level1_first
heatmap_level1_middle
heatmap_level1_last
heatmap_level2
heatmap_level2_first
heatmap_level2_middle
heatmap_level2_last
heatmap_level3
heatmap_level3_first
heatmap_level3_middle
heatmap_level3_last

barplot_level1
barplot_level1_first
barplot_level1_middle
barplot_level1_last
barplot_level2
barplot_level2_first
barplot_level2_middle
barplot_level2_last
barplot_level3
barplot_level3_first
barplot_level3_middle
barplot_level3_last
```
## Teddy's Comments
This is just some playing around with visualizing the accuracy data. Some more clean up will be needed and plotting data at specific instances (first encounter, last encounter etc). The plots below show mean accuracy across all instances (which is something we want to separate out). 
Think about doing some boxplots to see how the data is distributed for accuracy and reaction time data. The descriptives can be used to answer questions like: 
 - which items were most difficult for students?
 - what other patterns are salient in behavior?
 - How many trials does it take students, on average, to learn the associations? You can show change (increase) in accuracy across sessions and/or change (decrease) in reaction time across sessions. 
 - Do you have other questions that come in mind for you about this dat

## Original code
```{r}
dat4plot <- 
  dat %>% 
  select(given_response, cue_text, correct, level, fact_id) %>% 
  filter(level==1) %>% 
  mutate(cue_text=fct_reorder(cue_text,fact_id)) %>% 
    group_by(cue_text) %>% 
  summarise(mean_acc = mean(correct),
            n=n(), 
            se = sd(correct)/sqrt(n)) %>% 
  mutate(
      multiplier = as.numeric(str_extract(cue_text, "^\\d+")), # Extract the first number in `cue_text`
      multiplicand = as.numeric(str_extract(cue_text, "\\d+$")) # Extract the second number in `cue_text`
    ) %>% 
  mutate(cue_text=fct_reorder(cue_text,multiplicand)) 

dat4plot %>% 
  ggplot(aes(x=cue_text, y=mean_acc, ymax=mean_acc+se, ymin=mean_acc-se, fill=as.factor(multiplicand)))+
  geom_bar(stat = 'identity')+
  geom_errorbar()+
  theme(axis.text.x=element_text(angle = 90), 
        legend.position = 'top')

dat4plot %>% 
  ggplot(aes(x=multiplier, y = multiplicand, fill=mean_acc, label=round(mean_acc, 2))) +
  geom_tile() +
  geom_text()+
  scale_fill_gradient2(limits=c(0.7,1),
                       midpoint = 0.85,
                          low='blue',
                        mid="red",
                           high = 'yellow',
                          guide='colorbar',
                          aesthetics = 'fill',
                          breaks= seq(0.7,1,by=.01)
                       )

```
